DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life

### 2023-02-17 manuscript/manuscript.tex
remove figures from main text and finish answers to reviewer 1

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations\newline\indent Word count: 5480}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

Building a chronogram is not an easy task.
It requires obtaining and curating data to construct a phylogeny, selecting and placing appropriate calibrations on the phylogeny using independent age data points from the fossil record or other dated events, and inferring the full dated tree;
it also generally requires specialized biological training, taxonomic domain knowledge, and a non-negligible amount of research time, computational resources and funding.

Here we present the DateLife project which has the main goal of capturing age data from published chronograms, and making these data readily accessible to the community for reuse and reanalysis, for research, teaching, and science communication and policy.
DateLife's core software application is available as an R package (Sanchez-Reyes et al., 2022), and as an online Rshiny interactive website at www.datelife.org. It features key elements for scientific reproducibility, such as a versioned, open and fully public source database (McTavish et al., 2015) that stores data in a computer-readable format (Vos et al., 2012); automated and programmatic ways of accessing and downloading the data in a computer-readable format also (Stoltzfus et al., 2013); and methods to summarize and compare the data.

@@

DateLife's core software application consists of the R package \texttt{datelife}. Its current stable version -- v0.6.6, is available from The Comprehensive R Archive Network (CRAN) repository (Sanchez-Reyes et al., 2022), and relies on functionalities from various biological R packages:
ape (Paradis, Claude, \& Strimmer, 2004),
bold (Chamberlain, 2018),
geiger (Pennell et al., 2014),
paleotree (Bapst, 2012),
phyloch (Heibl, 2008),
phylocomr (Ooms \& Chamberlain, 2018),
phytools (Revell, 2012),
rotl (Michonneau, Brown, \& Winter, 2016), and
taxize (Chamberlain, 2018; Chamberlain \& Szöcs, 2013).
Figure \ref{fig:workflow} provides a graphical summary of the three main steps of the DateLife workflow: creating a search query, searching a database, and summarizing results from the search.

@@

At the time of writing of this manuscript (Jun 22, 2022),
DateLife's chronogram database latest version consist of 253 chronograms published in 187 different studies. It is curated from OpenTree's phylogenetic database, the Phylesystem, which constitutes an open source of expert and peer-reviewed phylogenetic knowledge with rich metadata (McTavish et al., 2015), which allows automatic and reproducible assembly of our chronogram database. Datelife's chronogram database is navigable as an R data object within the \texttt{datelife} R package.

@@

\emph{\textbf{Choosing a topology.--}}
DateLife requires a tree topology to summarize age data upon.
We recommend that users provide as input a tree topology from the literature, or one of their own making. If no topology is provided, DateLife automatically extracts one from the OpenTree synthetic tree, a phylogeny currently encompassing 2.3 million taxa across all life, assembled from 1, 239 published phylogenetic trees and OpenTree's unified Taxonomy, OTT (Open Tree Of Life et al., 2019).
Alternatively, DateLife can combine topologies from source chronograms using a supertree approach (Criscuolo, Berry, Douzery, \& Gascuel, 2006).
To do this, DateLife first identifies the source chronograms that form a grove, roughly, a sufficiently overlapping set of taxa between trees, by implementing definition 2.8 for n-overlap from Ané et al. (2009).
If the source chronograms do not form a grove, the supertree reconstruction will fail.
In rare cases, a group of trees can have multiple groves. By default, DateLife chooses the grove with the most taxa, however, the \enquote{criterion = trees} flag allows the user to choose the grove with the most trees instead.
The result is a single summary (or supertree) topology, that combines topologies from source chronograms in a grove.

\emph{\textbf{Applying secondary calibrations.--}}
Once a topology is chosen, DateLife applies the congruification method (Eastman, Harmon, \& Tank, 2013) that find nodes belonging to the same clade across source chronograms, and then extracts the corresponding node ages from patristic distance matrices stored as a \texttt{datelifeResult} object. Note that by definition, these matrices store total distance (time from tip to tip), assuming that the terminal taxa are coeval and occur at the present. Hence, node ages correspond to half the values stored in the \texttt{datelifeResult} matrices.
A table of congruified node ages that can be used as calibrations for a dating analysis is stored as a \texttt{congruifiedCalibrations} object.

@@

\emph{\textbf{Dating a tree topology.--}}
By default, DateLife implements the Branch Length Adjuster (BLADJ; Webb et al., 2008; Webb \& Donoghue, 2005) algorithm to obtain a fully dated topology.
BLADJ takes an initial topology and ignores branch length data if present. Next, it fixes ages on nodes with calibration data, and distributes time for nodes with no data evenly between calibrated nodes, minimizing age variance in the resulting chronogram (Webb et al., 2008). This method is useful for ecological analyses (Webb et al., 2008).
When there is conflict between ages of calibrated nodes, ignores node ages that are older than the age of a parent node.
BLADJ requires a root age estimate. If there is no information on the age of the root in the chronogram database, users can provide an estimate from the literature. If none is provided, DateLife will not return a dated topology and will provide a warning message along with suggestions on how the user can provide an age for the root so the dating analysis can be finished.

Alternatively, the user can choose to obtain a fully dated phylogeny
MrBayes (Huelsenbeck \& Ronquist, 2001; Ronquist \& Huelsenbeck, 2003) approach in DateLife uses the calibrations as priors on node ages.

The latter methods can assign dates using a birth-death model (all of them????), they require initial branch lengths and thus require more time and expertise to run, which makes BLADJ the fastest and most practical method to obtain a dated tree.

In the case the user can provide or obtain branch length data along with a tree topology, phylogenetic dating options supported in DateLife that incorporate branch length information from the input topology in combination with the secondary calibrations:
PATHd8 is a non-clock, rate-smoothing method (Britton et al., 2007) to date trees;
treePL (Smith \& O'Meara, 2012), is a semi-parametric, rate-smoothing, penalized likelihood dating method (Sanderson, 2002).

\emph{\textbf{Dating a tree with branch lengths--.}}
Topologies obtained from OpenTree and with the supertree approach described above lack branch length data.
Yet, phylogenetic dating using branch lengths data is the golden standard for phylogenetic dating analyses, but it is costly and requires a lot of human/expert curation.

A fast solution implemented in DateLife is as follows.
To estimate branch lengths proportional to substitution rates for these topologies, DateLife currently implements a simple algorithm. First, it mines the Barcode of Life Data System, BOLD (Ratnasingham \& Hebert, 2007) to obtain genetic markers for the input taxa.
Mined genetic sequences are aligned with MUSCLE (Edgar, 2004) (by default) or MAFFT (Katoh, Asimenos, \& Toh, 2009).

The BOLD sequence alignment is then used to obtain initial branch lengths with the accelerated transformation (ACCTRAN) parsimony algorithm, which resolves ambiguous character optimization, by assigning changes along branches of the tree as close to the root as possible (Agnarsson \& Miller, 2008). This results in older nodes (Forest et al., 2005).
Optionally, the likelihood of the tree topology, the alignment and the reconstructed branch lengths given different evolutionary models, is computed using functions from the \texttt{phangorn} package (Schliep, 2011).

Relative branch length information provides key data for phylogenetic dating, especially for nodes without secondary calibrations available. Yet, topologies without branch lengths can also be dated.

\emph{\textbf{Visualizing results.--}}
Finally, users can save all source and summary chronograms in formats that permit reuse and reanalysis (such as newick and the R \enquote{phylo} format), as well as visualize and compare results graphically, or construct their own graphs using DateLife's chronogram plot generation functions available from the R package \texttt{datelifeplot} (Sanchez-Reyes \& O'Meara, 2022).

@@

DateLife used the processed input names to search the local chronogram database and found 9 matching chronograms in 6 different studies (Fig. \ref{fig:figure2}B). Three studies matched five input names (Barker, Burns, Klicka, Lanyon, \& Lovette, 2015; Hedges, Marin, Suleski, Paymer, \& Kumar, 2015; Jetz, Thomas, Joy, Hartmann, \& Mooers, 2012), one study matched four input names (Hooper \& Price, 2017) and two studies matched two input names (Barker, Burns, Klicka, Lanyon, \& Lovette, 2013; Burns et al., 2014). No studies matched all input names. Together, source chronograms provide 28 unique age data points, covering all nodes on our chosen tree topology to date (Table \ref{tbl:table1}).

\vspace{3mm}

@@

\vspace{50mm}

@@

\hypertarget{primary-vs-secondary-calibrations}{%
\subsubsection{Primary vs Secondary calibrations}\label{primary-vs-secondary-calibrations}}

@@

The DateLife software is free and open source and it can be used through its R shiny web application at \url{http://www.datelife.org}, through the \texttt{datelife} R package, and through Phylotastic's project web portal \url{https://phylo.cs.nmsu.edu/}.
DateLife's web application is maintained using RStudio's shiny server and the shiny package open infrastructure, as well as Docker and OpenTree's infrastructure (dates.opentreeoflife.org/datelife).
\texttt{datelife}'s R package stable version is available for installation from the CRAN repository (\url{https://cran.r-project.org/package=datelife}) using the command \texttt{install.packages(pkgs\ =\ "datelife")} from within R.
Development versions are available from the GitHub repository (\url{https://github.com/phylotastic/datelife}) and can be installed using the command \texttt{devtools::install\_github("phylotastic/datelife")}.

@@

Supplementary material, including code, biological examples and benchmark results data files and online-only appendices, can be found in the Dryad data repository (\url{https://doi.org/10.5061/dryad.cnp5hqc6w}), as well as in the GitHub repositories used to develop the reproducible manuscript (\url{https://doi.org/10.5281/zenodo.7435094}), the biological examples (\url{https://doi.org/10.5281/zenodo.7435101}), and the software benchmark (\url{https://doi.org/10.5281/zenodo.7435106}).

@@

The DateLife project was born as a prototype tool aiming to provide these services, and was initially developed over a series of hackathons at the National Evolutionary Synthesis Center, NC, USA (Stoltzfus et al., 2013).
We thank colleagues from the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project. The team that assembled DateLife's first proof of concept: Tracy Heath, Jonathan Eastman, Peter Midford, Joseph Brown, Matt Pennell, Mike Alfaro, and Luke Harmon.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used to construct DateLife's chronogram database.
The many scientists who publish their chronograms in an open, reusable form, and the scientists who curate them for deposition in the Open Tree of Life repository.
The NSF for funding nearly all the above, in addition to the ABI grant that funded this project itself.

@@

@@

\leavevmode\hypertarget{ref-datelifeR}{}%
Sanchez-Reyes, L. L., O'Meara, B., Eastman, J., Heath, T., Wright, A., Schliep, K., \ldots{} Alfaro, M. (2022). \texttt{datelife}: Scientific Data on Time of Lineage Divergence for Your Taxa. \emph{R Package Release V0.6.2}. Retrieved from \url{https://doi.org/10.5281/zenodo.593938}

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations\newline\indent Word count: 5629}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

@@

Building a chronogram is not an easy task.
It requires obtaining and curating a homology hypothesis to construct a phylogeny, selecting and placing appropriate calibrations on the phylogeny using independent age data points from the fossil record or other dated events, and inferring a full dated tree. All of this entails specialized biological training, taxonomic domain knowledge, and a non-negligible amount of research time, computational resources and funding.

Here we present the DateLife project which has the main goal of extracting and exposing age data from published chronograms, making age data readily accessible to the wider community for reuse and reanalysis, research, teaching, and science communication and policy.
DateLife's core software application is available as an R package (Sanchez-Reyes et al., 2022), and as an online Rshiny interactive website at www.datelife.org. It features key elements for scientific reproducibility, such as a curated, versioned, open and fully public chronogram database (McTavish et al., 2015) that stores data in a computer-readable format (Vos et al., 2012); automated and programmatic ways of accessing and downloading the data in a computer-readable format also (Stoltzfus et al., 2013); and methods to summarize and compare the data.

@@

DateLife's core software application consists of the R package \texttt{datelife}. Its current stable version -- v0.6.6, is available from The Comprehensive R Archive Network (CRAN) repository (Sanchez-Reyes et al., 2022), and relies on functionalities from various biological R packages:
ape (Paradis, Claude, \& Strimmer, 2004),
bold (Chamberlain, 2018),
geiger (Pennell et al., 2014),
msa (Bodenhofer, Bonatesta, Horejš-Kainrath, \& Hochreiter, 2015),
paleotree (Bapst, 2012),
phyloch (Heibl, 2008),
phylocomr (Ooms \& Chamberlain, 2018),
phytools (Revell, 2012),
rotl (Michonneau, Brown, \& Winter, 2016), and
taxize (Chamberlain, 2018; Chamberlain \& Szöcs, 2013).
Figure \ref{fig:workflow} provides a graphical summary of the three main steps of the DateLife workflow: creating a search query, searching a database, and summarizing results from the search.

@@

At the time of writing of this manuscript
(Feb 20, 2023),
DateLife's chronogram database latest version consist of 253 chronograms published in 187 different studies. It is curated from OpenTree's phylogenetic database, the Phylesystem, which constitutes an open source of expert and peer-reviewed phylogenetic knowledge with rich metadata (McTavish et al., 2015), which allows automatic and reproducible assembly of our chronogram database. Datelife's chronogram database is navigable as an R data object within the \texttt{datelife} R package.

@@

\hypertarget{choosing-a-topology}{%
\subsection{Choosing a topology}\label{choosing-a-topology}}

DateLife requires a tree topology to summarize age data upon.
We recommend that users provide as input a tree topology from the literature, or one of their own making. If no topology is provided, DateLife automatically extracts one from the OpenTree synthetic tree, a phylogeny currently encompassing 2.3 million taxa across all life, assembled from 1, 239 published phylogenetic trees and OpenTree's unified Taxonomy, OTT (Open Tree Of Life et al., 2019).
Alternatively, DateLife can combine topologies from source chronograms using a supertree approach (Criscuolo, Berry, Douzery, \& Gascuel, 2006).
To do this, DateLife first identifies the source chronograms that form a grove, roughly, a sufficiently overlapping set of taxa between trees, by implementing definition 2.8 for n-overlap from Ané et al. (2009).
If the source chronograms do not form a grove, the supertree reconstruction will fail.
In rare cases, a group of trees can have multiple groves. By default, DateLife chooses the grove with the most taxa, however, the \enquote{criterion = trees} flag allows the user to choose the grove with the most trees instead.
The result is a single summary (or supertree) topology, that combines topologies from source chronograms in a grove.

\hypertarget{applying-secondary-calibrations}{%
\subsection{Applying secondary calibrations}\label{applying-secondary-calibrations}}

Once a topology is chosen, DateLife applies the congruification method (Eastman, Harmon, \& Tank, 2013) that find nodes belonging to the same clade across source chronograms, and then extracts the corresponding node ages from patristic distance matrices stored as a \texttt{datelifeResult} object. Note that by definition, these matrices store total distance (time from tip to tip), assuming that the terminal taxa are coeval and occur at the present. Hence, node ages correspond to half the values stored in the \texttt{datelifeResult} matrices.
A table of congruified node ages that can be used as calibrations for a dating analysis is stored as a \texttt{congruifiedCalibrations} object.

@@

\hypertarget{dating-a-tree-topology}{%
\subsection{Dating a tree topology}\label{dating-a-tree-topology}}

\hypertarget{with-no-branch-lengths}{%
\subsubsection{With no branch lengths}\label{with-no-branch-lengths}}

When producing or obtaining a tree with branch lengths for a group of interest is not possible, DateLife can date a topology without branch lengths, obtained from OpenTree or by implementing the supertree approach described above, by implementing the Branch Length Adjuster (BLADJ; Webb et al., 2008; Webb \& Donoghue, 2005) algorithm, which requires no initial branch lengths.
The algorithm starts by fixing ages for nodes with calibration data upon the given topology. Then, it distributes time for nodes with no data evenly between calibrated nodes, minimizing age variance in the resulting chronogram (Webb et al., 2008), which has proven useful for ecological analyses that require age data (Webb et al., 2008).
When there is conflict between ages of calibrated nodes, BLADJ ignores node ages that are older than the age of a parent node.
The BLADJ algorithm requires a root age estimate to run. If there is no information on the age of the root in the chronogram database, users can provide an estimate from the literature. If none is provided, DateLife will not return a dated topology and will provide a warning message along with suggestions on how the user can provide an age for the root so that a fully dated tree can be returned.

In the absence of genetic data, we consider that BLADJ is the most agnostic way to assign ages to nodes with no available data, as it does not require any assumptions on the underlying model of branch length distribution. It is however common practice in the literature to use a birth-death model to assign ages to nodes with no genetic data (Jetz, Thomas, Joy, Hartmann, \& Mooers, 2012; Rabosky et al., 2018; Smith \& Brown, 2018). To do so, DateLife implements MrBayes (Huelsenbeck \& Ronquist, 2001; Ronquist \& Huelsenbeck, 2003), using nodes with published age data as calibration priors on a fixed topology, a simple birth-death model with parameters that can be determined by the user, and no genetic data.

\hypertarget{dating-a-tree-with-branch-lengths}{%
\subsubsection{Dating a tree with branch lengths}\label{dating-a-tree-with-branch-lengths}}

Relative branch lengths provide key information for phylogenetic dating, specifically for nodes without any calibration data available.
While, using initial branch lengths data is the golden standard for phylogenetic dating analyses, producing such requires assembling and curating a homology hypothesis, and choosing and implementing a method for phylogenetic inference.
DateLife implements a workflow to streamline this process by leveraging on genetic data from the Barcode of Life Data System, BOLD (Ratnasingham \& Hebert, 2007) to obtain markers for input taxa.
By default, mined genetic sequences are aligned with MUSCLE (Edgar, 2004) using functions from the msa R package (Bodenhofer et al., 2015). Alternatively, sequences can be aligned with MAFFT (Katoh, Asimenos, \& Toh, 2009), using functions from the ape R package (Paradis et al., 2004).
The BOLD sequence alignment is then used to obtain initial branch lengths with the accelerated transformation (ACCTRAN) parsimony algorithm, which resolves ambiguous character optimization by assigning changes along branches of the tree as close to the root as possible (Agnarsson \& Miller, 2008), resulting in older nodes (Forest et al., 2005). The parsimony branch lengths are then optimized using Maximum Likelihood, given the alignment, the topology and a simple Jukes-Cantor model, producing a BOLD tree with branch lengths proportional to expected number of substitutions per site. Both parsimony and ML optimizations are done with functions from the \texttt{phangorn} package (Schliep, 2011).
The BOLD workflow is currently only available on DateLife's R package and not on the web application.

BOLD trees generated by DateLife, and any tree with branch lengths provided by a user can be analysed with phylogenetic dating methods supported in DateLife that incorporate branch length information from the input topology in combination with the secondary calibrations:
PATHd8, a non-clock, rate-smoothing method to date trees (Britton et al., 2007);
treePL (Smith \& O'Meara, 2012), a semi-parametric, rate-smoothing, penalized likelihood dating method (Sanderson, 2002);
and MRBayes (Huelsenbeck \& Ronquist, 2001; Ronquist \& Huelsenbeck, 2003), a Bayesian inference program implementing Markov chain Monte Carlo (MCMC) methods to estimate a posterior distribution of model parameters.

\hypertarget{visualizing-results}{%
\subsection{Visualizing results}\label{visualizing-results}}

Finally, users can save all source and summary chronograms in formats allowing for reuse and reanalysis, such as newick and the R \enquote{phylo} format. Input and summary chronograms can be visualized and compared graphically, and users can construct their own graphs using DateLife's chronogram plot generation functions available from the R package \texttt{datelifeplot} (Sanchez-Reyes \& O'Meara, 2022).

@@

DateLife used the processed input names to search the local chronogram database and found 9 matching chronograms in 6 different studies (Fig. \ref{fig:figure2}B). Three studies matched five input names (Barker, Burns, Klicka, Lanyon, \& Lovette, 2015; Hedges, Marin, Suleski, Paymer, \& Kumar, 2015; Jetz et al., 2012), one study matched four input names (Hooper \& Price, 2017) and two studies matched two input names (Barker, Burns, Klicka, Lanyon, \& Lovette, 2013; Burns et al., 2014). No studies matched all input names. Together, source chronograms provide 28 unique age data points, covering all nodes on our chosen tree topology to date (Table \ref{tbl:table1}).

@@

@@

\hypertarget{primary-vs-secondary-calibrations}{%
\subsection{Primary vs Secondary calibrations}\label{primary-vs-secondary-calibrations}}

@@

The DateLife software is free and open source. It can be used online through its R shiny web application hosted at \url{http://www.datelife.org}, and locally through the \texttt{datelife} R package, downloadable from Zenodo (\url{https://doi.org/10.5281/zenodo.593938}) and the CRAN repository (Sanchez-Reyes et al., 2022).
DateLife's web application is maintained using RStudio's shiny server and the shiny package open infrastructure, as well as Docker and OpenTree's infrastructure (dates.opentreeoflife.org/datelife).
\texttt{datelife}'s R package stable version is available for installation from CRAN repository using the command \texttt{install.packages(pkgs\ =\ "datelife")} from within R.
Development versions are available from DateLife's GitHub repository (\url{https://github.com/phylotastic/datelife}) and can be installed using the command \texttt{devtools::install\_github("phylotastic/datelife")}.

@@

Supplementary material, including code, biological examples, benchmark results, data files and online-only appendices, can be found in the Dryad data repository (\url{https://doi.org/10.5061/dryad.cnp5hqc6w}), as well as in the Zenodo stable repositories that host the reproducible manuscript (\url{https://doi.org/10.5281/zenodo.7435094}), the biological examples (\url{https://doi.org/10.5281/zenodo.7435101}), and the software benchmark (\url{https://doi.org/10.5281/zenodo.7435106}). The corresponding development versions hosted on GitHub can be found at \url{https://github.com/LunaSare/datelifeMS1\#readme}, \url{https://github.com/LunaSare/datelife_examples\#readme}, and \url{https://github.com/LunaSare/datelife_benchmark\#readme}.

@@

The DateLife project was born as a prototype tool aiming to provide the services describe in this paper, and was initially developed over a series of hackathons at the National Evolutionary Synthesis Center, NC, USA (Stoltzfus et al., 2013).
We thank colleagues from the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project. The team that assembled DateLife's first proof of concept: Tracy Heath, Jonathan Eastman, Peter Midford, Joseph Brown, Matt Pennell, Mike Alfaro, and Luke Harmon.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used to construct DateLife's chronogram database.
The many scientists who publish their chronograms in an open, reusable form, and the scientists who curate them for deposition in the Open Tree of Life repository.
The NSF for funding nearly all the above, in addition to the ABI grant that funded this project itself.

@@

\leavevmode\hypertarget{ref-bodenhofer2015msa}{}%
Bodenhofer, U., Bonatesta, E., Horejš-Kainrath, C., \& Hochreiter, S. (2015). Msa: An r package for multiple sequence alignment. \emph{Bioinformatics}, \emph{31}(24), 3997--3999.

@@

\leavevmode\hypertarget{ref-sanchez2022}{}%
Sanchez-Reyes, L. L., O'Meara, B., Eastman, J., Heath, T., Wright, A., Schliep, K., \ldots{} Alfaro, M. (2022). \texttt{datelife}: Scientific Data on Time of Lineage Divergence for Your Taxa. In \emph{R package version 0.6.6}. Retrieved from \href{https://CRAN.R-project.org/package=datelife\%20and\%20https://doi.org/10.5281/zenodo.593938}{https://CRAN.R-project.org/package=datelife and https://doi.org/10.5281/zenodo.593938}
